Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-18 ieee_paper.tex
bandwidth

In this experimental section, as a point of reference we provide some
results on a typical CPU architecture. This system is ...\todo{include
  a short description of the host.}
All the codes are compiled with icc xxx and -O3 optimization.

@@

When using the MIC card (Figure~\ref{fig:band_rw_MIC}) the {\tt
  read\_cpp} implementation reaches a performance of 160GB/s
\todo{gordon check exact number} which is better than the {\tt read}
implementation which only achieves 145GB/s \todo{and that one}. This
is explained by the compiler adding loop unrolling and software
prefetching in the assembly code in the {\tt read\_cpp} case which is
not added in the {\tt read} case. On the other hand, the {\tt write}
case reaches a performance of 155GB/s\todo{and that one} while the
{\tt write\_cpp} implementation never passes 55GB/s. This difference
is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction
which bypasses the Read For Ownership protocol and allows the writes
to be performed in any order. The compiler can not use that
instruction on its own because using would break the memory convention
useful for some parallel algorithms. The {\tt read/write} benchmark
performs at 140GB/s while the {\tt read/write cpp} benchmark performs
at about 90GB/s.

@@

@@

%<<<<<<< HEAD
%\begin{figure}
%  \centering
%  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
%
%  \caption{Performance on a MIC of derivative computation for RBFFD
%    with 884,736 nodes distributed quasi-randomly in a cube. Single
%    derivative of a single function (base case, triangle) and four
%    derivatives of four functions (multi case, circles). No bandwidth
%    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
%\end{figure}
%
%\begin{figure*}
%\centering
%\subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}} %\label{fig:ratio_best}}
%\subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}}%\label{fig:ratio_worst}}
%\subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}}%\label{fig:ratio_nocache}}
%\caption{Ratio of flops to bytes; (a) simple count of flops and bytes that require transfer from memory;(b) best and (c) worst
%case taking into account that bringing one element of $x$ into memory brings along the remainder of the cache line.}
%\label{fig:ratio_bytes_flops}
%\end{figure*}
%=======
%\begin{figure}
%  \centering
%  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
%
%  \caption{Performance on a MIC of derivative computation for RBFFD
%    with 884,736 nodes distributed quasi-randomly in a cube. Single
%    derivative of a single function (base case, triangle) and four
%    derivatives of four functions (multi case, circles). No bandwidth
%    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
%\end{figure}
%
%>>>>>>> 619115cea6da9466f09a3b01456eacf7991606e2

>>>>

In this experimental section, as a point of reference we provide some
results on a typical CPU architecture. This system is ...\todo{include
  a short description of the host.}  All the codes are compiled with
icc xxx and -O3 optimization. The MIC codes use the IMCI instructions
while the CPU codes use AVX instructions.

@@

When using the MIC card (Figure~\ref{fig:band_rw_MIC}) the {\tt
  read\_cpp} implementation reaches a performance of 160GB/s
\todo{gordon check exact number} which is better than the {\tt read}
implementation which only achieves 145GB/s \todo{and that one}. This
is explained by the compiler adding loop unrolling and software
prefetching in the assembly code in the {\tt read\_cpp} case which is
not added in the {\tt read} case. On the other hand, the {\tt write}
case reaches a performance of 155GB/s\todo{and that one} while the
{\tt write\_cpp} implementation never passes 55GB/s. This difference
is due to the use of the {\tt \_mm512\_store\_nrngo\_ps} instruction
which bypasses the Read For Ownership protocol and allows the writes
to be performed in any order. The compiler can not use this
instruction on its own because using would break the memory convention
useful for some parallel algorithms. The {\tt read/write} benchmark
performs at 140GB/s while the {\tt read/write cpp} benchmark performs
at about 90GB/s.

Looking at the results on the host in Figure~\ref{fig:band_rw_CPU}
show no significant difference between manual vectorization and the
compiler generated one (there is no store nrngo on the host). It also
shows quite stable performance with about 70GB/s in the {\tt read}
benchmark, about 40GB/s in the {\tt write} benchmark and about 49 in
the {\tt read write} benchmark. The bandwidth achieved on MIC is about
2 to 3 times faster than the bandwidth achieved on the CPU. The
kernels we consider being memory bound, we expect MIC to be 2 to 3
times faster.

Note that SpMV kernels will likely use only simple read and simple
writes on the MIC architecture. We also benchmark the performance we
can achieve using {\tt gather} and {\tt unpack} operations and we
present the results in Figure~\ref{fig:band_gather}. The {\tt unpack}
instruction can reach a performance of 140GB/s. (There is no easy way
to implement the equivalent of the {\tt unpack} operation without
using explicit registers.) The performance of the {\tt gather}
operations depends in which order the array is read. If the indices
are ordered ({\tt compact}) then a performance of 140GB/s is
reached. If the indices are completely randomized, the performance
drops to 2GB/s\todo{correct?}. We can see that implementing the
indirection mechanism without vector instructions makes the
performance drop significantly. (In the {\tt compact} case, it never
passes 100GB/s and is close to 50GB/s most of the time.)

@@

To summarize, using vector instruction seems mandatory to reach the
highest bandwidth on the MIC architecture. For the instructions used
in a typical SpMV computation, one can expect a bandwidth between
140GB/s to 160GB/s. That is why we set in our estimation the best
achievable bandwidth to 150GB/s despite carefully crafted code to
maximize bandwidth can reach higher
values~\cite{Saule13-ARXIV}. Comparing to a CPU architecture, we
expect MIC to be 2 to 3 times faster.

@@
